### 2018-04-04 ms/ms_probaweb.tex
final edits and compilation

Community ecology is \textit{the study of the interactions that determine the
distribution and abundance of organisms} \citep{Krebs2001}. Despite a general
consensus on this definition \citep{Scheiner2007a}, research on variation in
community structure (beta-diversity) has mostly focused on the spatial and
temporal turnover of species composition \citep{Anderson2011}. Such research
usually neglects variation in the way species interact with each other,
despite accumulating empirical evidence that is a major source of
diversity \citep{Poisot2015a}. Given this omission, it is perhaps not
surprising that biogeographers are still struggling to establish whether
interactions actually impact the distribution of species at large spatial
scales \citep{Kissling2012}. An interaction is conceived as the direct effect
of the action of one species on the demography on another species; whether it
scales up and impact the (co-)distribution of species remains matter of debate
at the moment \citep{Wisz2012, GonzalezSalazar2013, Cazelles2016b,
Harris2016, Godsoe2017}. Treating interactions as fixed events nonetheless
neglects a large part of the complexity of empirical communities, and will
most likely deliver biased metrics \citep{Poisot2015c}. Recent attempts
at accounting for interactions in species distribution models
\citep{Pollock2014, Pelissier2013, Ovaskainen2017} have brought some
methodological advances, but are not sufficient for two reasons. First, these
techniques are still based on a `species-based' approach to communities, where
interactions are merely treated as fixed covariates affecting distribution.
Second, they failed to provide a conceptual step forward, both in their
treatment of interactions and in the quality of the predictions they make.

@@

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is a need to bring the two together. Here, we
offer an integrated approach to do so, adopting the view that community
structure is best represented as a network of ecological interactions. Based
on this idea, we propose a description of the ecological niche that
integrates the effect of the environment on species distribution and on the
ecological interactions among them. Building on this concept, we develop a
quantitative theory to explain turnover of interactions in space and time. We
first present the concept and then formalize it mathematically, using a
probabilistic model to represent the sampling of the regional pool of
interactions. At the level of species pairs, the statistical approach could be
conceived as an interaction distribution model. At the community level, the
approach provides a likelihood-based method to compare different hypotheses of
network turnover. As an illustrative example, we apply this novel framework to
a large data set on host–parasite interactions across Europe and find that two
aspects of the environment (temperature and precipitation) exert a strong
imprint on species co-occurrence, but not on species interactions. The network
structure changes systematically across the latitudinal gradient, with a peak
of connectance at intermediate latitudes.

@@

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community
\citep{Gotzenberger2012}?} This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Fig. 1. The regional pool of
interactions, the \emph{metaweb}, represents potential interactions among all
species that could be found in a given area. In this particular case, there
are 275 nodes, and 1173 links among the plants (52 nodes), herbivores (96
nodes), and parasitoids (127 nodes) from Northern Europe. An instance of a
local community is also illustrated, with 45 nodes and 93 interactions. Only
55.0\% of all potential interactions (plant-herbivore or herbivore parasitoid
combinations) are realized in this local network, revealing the stochastic
nature of ecological interactions. Our objective here is to provide theory to
explain the sampling of the regional pool of interactions, along with a
quantitative method to predict it. The problem could be formalized
sequentially by understanding first why only a fraction of the species co-occur locally and second why these species do or do not interact.

@@

Interestingly, the food web literature also has its own "niche model" to
position a species in a community \citep{Williams2000a}, generalized later to
other types of interaction networks \citep{Eklof2013}. The niche of a species
in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over three to five different niche axes \citep{Williams2000a,
Eklof2013}. The niche model of food web structure and its variants have
successfully explained the complexity of a variety of networks, from food webs
to plant–pollinator systems \citep{Allesina2008, Williams2010, Eklof2013}.
This conceptual framework is, however, limited to local communities, and does
not provide any explanation for the turnover of network structure along
environmental gradients.

@@

We now turn to an illustration of our theory with the analysis of an empirical
dataset of host–parasite networks sampled throughout the south–north
environmental gradient in continental Europe \citep{Kopelke2017}. Our
objective here is to illustrate potential applications of the approach
outlined above, rather than to specifically describe the biogeography
of this particular community. The focal system consists of local food webs of
willows (genus \textit{Salix}), their galling insects, and the natural enemies
(parasitoids and inquilines) of gallers. We ask:
i) how much does network structure vary across the gradient, and ii) what is
the primary driver of network turnover across the gradient?

@@

Local realizations of the willow–galler–parasitoid network were reconstructed
from community samples collected between 1982 and 2010. During this period,
willow galls were collected at 370 sites across Central and Northern Europe.
In total, 52, 96 and 127 \textit{Salix}, galler and parasitoid and inquiline
taxa were distinguished, respectively. The strength of this dataset is that
observations were observed in situ, rather than inferred from expert knowledge
or other sources of information, thereby allowing the analysis of their
spatial variation. The drawback is that, because many species are rare, some
of the pairs of species may have been observed co-occurring only a few times,
or never. As a consequence, despite the extent of the sampling, there is significant uncertainty in the quantification of some of the links. The above described methodology explicitly aims at describing this uncertainty, and research should be done to develop methods to reduce it (see Conclusion).

@@

Despite the extensive sampling, many pairs of species were observed to co-occur only a few times. This made it difficult to evaluate interaction
probabilities with any reasonable confidence. Thus, we start with an
example of a single pair of species selected because of its high number of co-occurrences ($N_{ij}=38$): the leaf folder \textit{Phyllocolpa prussica} and
the parasidoid \textit{Chrysocharis elongata}. These two fairly abundant
species were observed $N_i=49$ and $N_j=121$ times, respectively, across the
370 sites, and they were found to interact with a marginal probability
$P(L_{ij})=0.55$, which means they interacted at 21 different locations. Here,
a comparison of model fit (Table 1) reveals that conditioning the interaction
probability on local environmental conditions adds no explanatory power beyond
a model assuming the same probability of interaction anywhere in space (Model 1
vs Model 2). Moreover, when the two species co-occur, the occurrence of the
interaction was insensitive to the environment (Model 2 vs Model 3).
Alternatively, climatic variables significantly impacted co-occurrence (Model
3 vs Model 4). The neutral model performed worse than the non-random co-occurrence model (Model 3 vs Model 6). The full model revealed that the
greatest interaction probability occurred at intermediate temperature and
precipitation, simply because this is where the two species most frequently
co-occur (Fig. 3). The probabilities of co-occurrence and interaction can be
represented in space, where we found that the highest interaction probability
occurred in Central Europe (Fig. 4).

We evaluated each model for all pairs of species in order to better understand
the large-scale drivers of network turnover. The results were highly
consistent among trophic levels (\textit{Salix}–gallers and
gallers–parasitoids; Table 2), despite large variability in the fit
of the models to the different pairs of species. This variability was
particularly important for the models of interactions given co-occurrence.
Across all pairs of species, the conditional representation of interactions
performed better than the marginal one (Model 1 vs Model 2); that is,
interactions did not occur systematically whenever the two species were found
co-occurring. Hence, in addition to species turnover, the stochastic nature of
interactions contributes to network variability. In total, we recorded 1,173
pairs of interactions, only 290 of which occurred more than five times. Out of
these 290 interactions, 143 were systematically detected whenever the two
species co-occurred. In the instances when species co-occurred, the two
environmental variables considered proved relatively poor predictors of their
interactions (Model 2 vs Model 3). Not surprisingly, for both types of
interactions (\textit{Salix}–galler and galler–parasitoid), the log-likelihood
increased when the environment was considered. However, the extra number of
parameters exceeded the gain in log-likelihood and inflated AIC. Therefore,
the most parsimonious model excluded the effect of the environment. On the
basis of log-likelihood only, co-occurrence was non-neutral for both
\textit{Salix}–galler and galler–parasitoid interactions. Thus, according to
AIC, the best model was the one of non-random co-occurrence (Model 3 vs Model
6) for both types of interactions.

@@

Applying the framework to our large data set on host–parasite interactions
across Europe revealed key features in the interaction between \textit{Salix}
taxa, their herbivores, and the natural enemies of these herbivores.
Consistent with a general increase in the diversity of \textit{Salix} towards
boreal areas \citep{Cronk2015}, overall species richness of the networks
increased towards the north. The distribution of \textit{Salix} species
richness largely matched those of gallers and parasitoids. These observations
within Europe are also matched by the ones found at a global scale for
\textit{Salix} \citep{Argus1997, Cronk2015, Wu2015} and sawflies
\citep{Kouki1994, Kouki1999}. Species richness in a common group of parasitic
wasps, the Ichneumonidae, was originally presumed to show a similar "reversed
latitudinal gradient", but this observation has been recently challenged by
findings of rather high ichneumonid diversity in the tropics
\citep{Veijalainen2013}. Nevertheless, ichneumonid subfamilies specifically
associated with sawflies (Ctenopelmatinae, Tryphoninae) are clearly less
diverse in the south.

@@

\textbf{Mapping the distribution of species richness, the number of links and
connectance across Europe}. The representation is based on predictions from
Model 3 (see Table 2) using climate data extracted from WorldClim over a 1km x 1km grid. Species richness is obtained by summation of individual
occurrence probabilities, and link density by summation of interaction
probabilities. Colour gradient ranges from low values (dark blue) to high
values (yellow).

@@

\begin{figure}[ht!]
\centering\includegraphics[width=0.4\textwidth]{figures/mw_sampling}
\end{figure}

>>>>

Community ecology is \textit{the study of the interactions that determine the
distribution and abundance of organisms} \citep{Krebs2001}. Despite a general
consensus on this definition \citep{Scheiner2007a}, research on variation in
community structure (beta-diversity) has mostly focused on the spatial and
temporal turnover of species composition \citep{Anderson2011}. Such research
usually neglects variation in the way species interact with each other,
despite accumulating empirical evidence that is a major source of
diversity \citep{Poisot2015a}. Given this omission, it is perhaps not
surprising that biogeographers are still struggling to establish whether
interactions actually impacts the distribution of species at large spatial
scales \citep{Kissling2012}. An interaction is conceived as the direct effect
of the action of one species on the demography on another species; whether it
scales up and impact the (co-)distribution of species remains matter of debate
at the moment \citep{Wisz2012, GonzalezSalasar2013, Cazelles2016b,
Harris2016, Godsoe2017}. Treating interactions as fixed events nonetheless
neglects a large part of the complexity of empirical communities, and will
most likely deliver biased metrics \citep{Poisot2015c}. Recent attempts
at accounting for interactions in species distribution models
\citep{Pollock2014, Pelissier2013, Ovaskainen2017} have brought some
methodological advances, but are not sufficient for two reasons. First, these
techniques are still based on a `species-based' approach to communities, where
interactions are merely treated as fixed covariates affecting distribution.
Second, they fail to provide a conceptual step forward, both in their
treatment of interactions and in the quality of the predictions they make.

@@

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is a need to bring the two together. Here, we
offer an integrated approach to do so, adopting the view that community
structure is best represented as a network of ecological interactions. Based
on this idea, we propose a description of the ecological niche that
integrates the effect of the environment on species distribution and on the
ecological interactions among them. Building on this concept, we develop a
quantitative theory to explain turnover of interactions in space and time. We
first present the concept and then formalize it mathematically, using a
probabilistic model to represent the sampling of the regional pool of
interactions. At the level of species pairs, the statistical approach could be
conceived as an interaction distribution model. At the community level, the
approach provides a likelihood-based method to compare different hypotheses of
network turnover. As an illustrative example, we apply this novel framework to
a large data set on host-parasite interactions across Europe and find that two
aspects of the environment (temperature and precipitation) exert a strong
imprint on species co-occurrence, but not on species interactions. The network
structure changes systematically across the latitudinal gradient, with a peak
of connectance at intermediate latitudes.

@@

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community}
\citep{Gotzenberger2012}? This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Fig. 1. The regional pool of
interactions, the \emph{metaweb}, represents potential interactions among all
species that could be found in a given area. In this particular case, there
are 275 nodes, and 1173 links among the plants (52 nodes), herbivores (96
nodes), and parasitoids (127 nodes) from Northern Europe. An instance of a
local community is also illustrated, with 45 nodes and 93 interactions. Only
55.0\% of all potential interactions (plant-herbivore or herbivore parasitoid
combinations) are realized in this local network, revealing the stochastic
nature of ecological interactions. Our objective here is to provide theory to
explain the sampling of the regional pool of interactions, along with a
quantitative method to predict it. The problem could be formalized
sequentially by understanding first why only a fraction of the species co-occur locally and second why these species do or do not interact.

@@

Interestingly, the food web literature also has its own "niche model" to
position a species in a community \citep{Williams2000a}, generalized later to
other types of interaction networks \citep{Eklof2013}. The niche of a species
in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over three to five different niche axes \citep{Williams2000a,
Eklof2013}. The niche model of food web structure and its variants have
successfully explained the complexity of a variety of networks, from food webs
to plant-pollinator systems \citep{Allesina2008, Williams2010, Eklof2013}.
This conceptual framework is, however, limited to local communities, and does
not provide any explanation for the turnover of network structure along
environmental gradients.

@@

We now turn to an illustration of our theory with the analysis of an empirical
dataset of host-parasite networks sampled throughout the south-north
environmental gradient in continental Europe \citep{Kopelke2017}. Our
objective here is to illustrate potential applications of the approach
outlined above, rather than to specifically describe the biogeography
of this particular community. The focal system consists of local food webs of
willows (genus \textit{Salix}), their galling insects, and the natural enemies
(parasitoids and inquilines) of gallers. We ask:
i) how much does network structure vary across the gradient, and ii) what is
the primary driver of network turnover across the gradient?

@@

Local realizations of the willow-galler-parasitoid network were reconstructed
from community samples collected between 1982 and 2010. During this period,
willow galls were collected at 370 sites across Central and Northern Europe.
In total, 52, 96 and 127 \textit{Salix}, galler and parasitoid and inquiline
taxa were distinguished, respectively. The strength of this dataset is that
observations were observed in situ, rather than inferred from expert knowledge
or other sources of information, thereby allowing the analysis of their
spatial variation. The drawback is that, because many species are rare, some
of the pairs of species may have been observed co-occurring only a few times,
or never. As a consequence, despite the extent of the sampling, there is significant uncertainty in the quantification of some of the links. The above described methodology explicitly aims at describing this uncertainty, and research should be done to develop methods to reduce it (see Conclusion).

@@

Despite the extensive sampling, many pairs of species were observed to co-occur only a few times. This made it difficult to evaluate interaction
probabilities with any reasonable confidence. Thus, we start with an
example of a single pair of species selected because of its high number of co-occurrences ($N_{ij}=38$): the leaf folder \textit{Phyllocolpa prussica} and
the parasidoid \textit{Chrysocharis elongata}. These two fairly abundant
species were observed $N_i=49$ and $N_j=121$ times, respectively, across the
370 sites, and they were found to interact with a marginal probability
$P(L_{ij})=0.55$, which means they interacted at 21 different locations. Here,
a comparison of model fit (Table 1) reveals that conditioning the interaction
probability on local environmental conditions adds no explanatory power beyond
a model assuming the same probability of interaction anywhere in space (Model 1
vs Model 2). Moreover, when the two species co-occur, the occurrence of the
interaction was insensitive to the environment (Model 2 vs Model 3).
Alternatively, climatic variables significantly impacted co-occurrence (Model
3 vs Model 4). The neutral model performed worse than the non-random co-occurrence model (Model 3 vs Model 6). The full model revealed that the
greatest interaction probability occurred at intermediate temperature and
precipitation, simply because this is where the two species most frequently
co-occur (Fig. 3). The probabilities of co-occurrence and interaction can be
represented in space, where we find that the highest interaction probability
occurred in Central Europe (Fig. 4).

We evaluated each model for all pairs of species in order to better understand
the large-scale drivers of network turnover. The results were highly
consistent among trophic levels (\textit{Salix}-gallers and
gallers-parasitoids; Table 2), despite large variability in the fit
of the models to the different pairs of species. This variability was
particularly important for the models of interactions given co-occurrence.
Across all pairs of species, the conditional representation of interactions
performed better than the marginal one (Model 1 vs Model 2); that is,
interactions did not occur systematically whenever the two species were found
co-occurring. Hence, in addition to species turnover, the stochastic nature of
interactions contributes to network variability. In total, we recorded 1,173
pairs of interactions, only 290 of which occurred more than five times. Out of
these 290 interactions, 143 were systematically detected whenever the two
species co-occurred. In the instances when species co-occurred, the two
environmental variables considered proved relatively poor predictors of their
interactions (Model 2 vs Model 3). Not surprisingly, for both types of
interactions (\textit{Salix}-galler and galler-parasitoid), the log-likelihood
increased when the environment was considered. However, the extra number of
parameters exceeded the gain in log-likelihood and inflated AIC. Therefore,
the most parsimonious model excluded the effect of the environment. On the
basis of log-likelihood only, co-occurrence was non-neutral for both
\textit{Salix}-galler and galler-parasitoid interactions. Thus, according to
AIC, the best model was the one of non-random co-occurrence (Model 3 vs Model
6) for both types of interactions.

@@

Applying the framework to our large data set on host-parasite interactions
across Europe revealed key features in the interaction between \textit{Salix}
taxa, their herbivores, and the natural enemies of these herbivores.
Consistent with a general increase in the diversity of \textit{Salix} towards
boreal areas \citep{Cronk2015}, overall species richness of the networks
increased towards the north. The distribution of \textit{Salix} species
richness largely matched those of gallers and parasitoids. These observations
within Europe are also matched by the ones found at a global scale for
\textit{Salix} \citep{Argus1997, Cronk2015, Wu2015} and sawflies
\citep{Kouki1994, Kouki1999}. Species richness in a common group of parasitic
wasps, the Ichneumonidae, was originally presumed to show a similar "reversed
latitudinal gradient", but this observation has been recently challenged by
findings of rather high ichneumonid diversity in the tropics
\citep{Veijalainen2013}. Nevertheless, ichneumonid subfamilies specifically
associated with sawflies (Ctenopelmatinae, Tryphoninae) are clearly less
diverse in the south.

@@

\textbf{Mapping the distribution of species richness, the number of links and
connectance across Europe}. The representation is based on predictions from
Model 3 (see Table 2) using climate data extracted from WorldClim over a 1km x 1km grid. Species richness is obtained by summation of individual
occurrence probabilities, and link density by summation of interaction
probabilities. Colowr gradient ranges from low values (dark blue) to high
values (yellow).

@@

\begin{figure}[ht!]
\centering\includegraphics[width=1\textwidth]{figures/mw_sampling}
\end{figure}
